---
title: 'Critical Role Series: Attendance'
author: Afton Coombs
date: '2020-05-22'
slug: critical-role-series-attendance
categories:
  - R
  - analysis
  - d&d
  - criticalrole
tags:
  - R Markdown
---

```{r setup, echo = FALSE, include = FALSE, message = FALSE}
library(magrittr)
```

## Introduction

This is a short exploratory data analysis of attendance on the web series Critical Role, in which 
voice actors play Dungeons and Dragoooooooons! It is a fantastic show, 
*and* some the amazing people at <a href = "https://www.critrolestats.com/stats-wm">
Crit Role Stats</a> have been tracking <a href = "https://twitter.com/CritRoleStats/">"monster stats, dice rolls, and anything else that can be quantified on Critical Role"</a>.

Critical Role is currently on hiatus due to COVID-19, so to fill the usual time slot on Thursdays, I've 
been having some fun with <a href = "https://www.critrolestats.com/stats-wm">
Critical Role Stats</a> data, mostly for the purposes of data visualization. This first one is about actor attendance on Critical Role.

## Background Information
During this season, Ashley Johnson (playing Yasha) had a conflicting 
role acting on a TV series, and was absent for several episodes, 
Laura Bailey (Jester) and Travis Willingham (Fjord) had a baby (yay!) and was also absent for a few 
episodes, and Taliesin and Liam had a smattering of absences 
for various reasons.

## The Data
Attendance data downloaded from a <a href = "https://docs.google.com/spreadsheets/d/1E1DfdXJVu9UpGNG29JMHT3ovk8Ol_UTzol40DMzz-rw/edit#gid=1399380767">Crit Role Stats Google sheet</a>.

```{r load-data}
attendance_data <- 
  read.csv("data/CR Attendance.csv",
           stringsAsFactors = FALSE)

head(attendance_data)  
```

With a little `dplyr` manipulation, we can add numeric episode values and a Date class column, as well
pivot the data into a long format which will be more useful for plotting. The column `value` here is a binary 
variable, indicating if an actor was (`1`) or wasn't (`0`) present for a particular episode.

I also converted the `character` type column `Actor` to a `factor` type, ordering by the seating
order of the show top left to bottom right. This will let us put the graph in the right order, when
we get to it. Because I want Travis to be top of the graph, and Ashley bottom of the graph, Travis must have the highest
factor value and Ashley must have the lowest, so the factor levels are set beginning with Ashley.

```{r data-munge}
attendance_data_long <-
  attendance_data %>%
  dplyr::mutate(Episode.Num = as.numeric(gsub("C2E", "", Episode)),
                Date = as.Date(Air.Date, format = "%m/%d/%y")) %>%
  tidyr::pivot_longer(cols = c("Laura", 
                               "Liam", 
                               "Marisha", 
                               "Taliesin", 
                               "Travis", 
                               "Sam", 
                               "Ashley"),
                      names_to = "Actor") %>%
  tibble::as_tibble() %>%
  dplyr::mutate(value = as.factor(value),
                Actor = factor(Actor,
                               levels = c("Ashley",
                                          "Taliesin",
                                          "Laura",
                                          "Sam",
                                          "Liam",
                                          "Marisha",
                                          "Travis")))

head(attendance_data_long)
```

## Plotting
It's fairly straightforward to plot the attendance data using a dark color marking "present" and a light
color marking "absent." 

What do we see here? Well, Ashley missed a bit while working on another TV show. Travis and Laura were absent 
on the same days starting around episode 25, when they had a baby. Sam and Marisha never missed an episode.

```{r simple-plot}
ggplot2::ggplot(data = attendance_data_long,
                ggplot2::aes(x = Episode.Num,
                             y = Actor,
                             fill = factor(value))) +
  ggplot2::geom_tile(color = "black") +
  ggthemes::theme_hc() +
  ggplot2::scale_color_manual(values = c("lightgray", "gray40"),
                              aesthetics = c("fill")) +
  ggplot2:: scale_x_continuous(breaks = seq(1, 94, 1),
                               labels = paste0("Ep. ", seq(1, 94, 1))) +
  ggplot2::theme(legend.position = "none",
                 axis.text.x = ggplot2::element_text(size = 5,
                                            angle = 90),
                 axis.ticks = ggplot2::element_blank(),
                 axis.title.x = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank())
```

To really make this pop, though, it would be nice to assign to each actor a color paletter reflective of their
character. For example, Marisha's character Beau belongs to the Cobalt Soul and wears blue robes, so a blue palette would be nice for her. Sam's character Nott is a goblin thief, so a green color palette would work. Each
color palette should have two colors, a dark color for "present" and a light color for "absent", so 7 actors * 2 values = 14 levels. Actually, Taliesin played two characters, Mollymauk until his death in episode 26, and Caduceus beginning
in episode 28, so we need 16 levels to assign a different color palette to Taliesin before and after the 26th.

These are hex codes for colors I thought looked right for each character, in a named vector where each name is one of the actors x attendance possibilities. I ordered the vector in the seating order of the show, with
Travis at top left and Ashley at bottom right.
This will let us use these colors as a palette when we get to the plot. 

```{r color-palette}
discrete_mighty_nein_colors <-
  c("Travis0" = as.character("#cadabe"),
    "Travis1" = as.character("#78a05a"),
    "Marisha0" = as.character("#9ebbfa"),
    "Marisha1" = as.character("#072d7f"),
    "Liam0" = as.character("#f3bea5"),
    "Liam1" = as.character("#b36817"),
    "Sam0" = as.character("#d8dabe"),
    "Sam1" = as.character("#525530"),
    "Laura0" = as.character("#b8c9e0"),
    "Laura1" = as.character("#5882b8"),
    "Taliesin0" = as.character("#cabcdc"),
    "Taliesin1" = as.character("#5b417d"),
    "Taliesin3" = as.character("#dfb9ca"),
    "Taliesin4" = as.character("#c47e9d"),
    "Ashley0" = as.character("#cccbcd"),
    "Ashley1" = as.character("#535154"))

discrete_mighty_nein_colors
```

We can plot the colors alone, just to get a sense of what they look like.

```{r test-palette}
ggplot2::ggplot(data = as.data.frame(discrete_mighty_nein_colors) %>% tibble::rownames_to_column(),
                mapping = ggplot2::aes(x = rowname, y = 1, fill = rowname)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::scale_color_manual(values = discrete_mighty_nein_colors,
                              aesthetics = c("fill")) +
  ggthemes::theme_hc() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90),
                 axis.title.x = ggplot2::element_blank(),
                 axis.ticks = ggplot2::element_blank(),
                 axis.text.y = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank(),
                 legend.position = "none")
```

You might be wondering why the labels are so flat -- e.g. "Taliesin0" and "Taliesin1", instead of a two-dimensional
structure that has "Taliesin" in one column and 1 or 0 in the other, like our attendance data already has.

Well, things gets tricky, because for `ggplot2` to give each character and their attendance value a different
color in the plot, each combination needs to be its own factor level. This can be achieved in a somewhat hacky way by
concatenating the actor name and their attendance value, e.g. "Taliesin0" for an episode in which Taliesin was absent.
If anyone knows a better way of achieving this outcome -- please let me know!

To use these colors, we need a new column in the data which has one factor level per actor and attendance combo.

```{r hackify-the-data}
attendance_data_long_with_factor <-
  attendance_data_long %>%
  dplyr::mutate(att_value = ifelse(test = Episode.Num > 27 & Actor == "Taliesin",
                                   yes = paste0(Actor, as.numeric(value) + 2),
                                   no = paste0(Actor, value))) %>%
  dplyr::mutate(att_value = factor(att_value,
                                   levels = names(discrete_mighty_nein_colors)))
```

```{r colorful-plot}
ggplot2::ggplot(data = attendance_data_long_with_factor,
                   ggplot2::aes(x = Episode.Num,
                                y = Actor,
                                fill = att_value)) +
  ggplot2::geom_tile() +
  ggthemes::theme_hc() +
  ggplot2::scale_color_manual(values = discrete_mighty_nein_colors,
                              aesthetics = c("fill")) +
  ggplot2:: scale_x_continuous(breaks = seq(1, 94, 1),
                               labels = paste0("Ep. ", seq(1, 94, 1))) +
  ggplot2::theme(legend.position = "none",
                 axis.text.x = ggplot2::element_text(size = 5,
                                            angle = 90),
                 axis.ticks = ggplot2::element_blank(),
                 axis.title.x = ggplot2::element_blank(),
                 axis.title.y = ggplot2::element_blank())
```