---
title: "Critical Role EDA"
author: "Afton Coombs"
date: "12/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
library(googlesheets4)
library(tidyr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(colorspace)
library(GetoptLong)
library(scater)
source("colors.R")
library(ggthemes)
library(cowplot)
library(colorspace)
library(colorblindr)
library(tidyverse)
library(network)
# library(igraph)
library(tidygraph)
library(ggraph)
library(png)
library(cowplot)
library(stringr)
library(tm)
library(SnowballC)
library(wordcloud)
```

## Introduction

This is an EDA of attendance on the web series Critical Role, in which 
voice actors play Dungeons and Dragoooooooons! It is a fantastic show, 
*and* some the amazing fans at <a href = "https://www.critrolestats.com/stats-wm">
Critical Role Stats</a> have been tracking <a href = "https://twitter.com/CritRoleStats/">"monster stats, dice rolls, and anything else that can be quantified on Critical Role"<a/>.

During this season, Ashley Johnson (playing Yasha) had a conflicting 
role acting on a TV series, and was absent for several episodes, 
Laura Bailey (Jester) had a baby (yay!) and was also absent for a few 
episodes, and Travis, Taliesin, and Liam had a smattering of absences 
for various reasons. On some occasions, the storyline was adapted to make
the character absent from the gameplay to correspond to the actor's 
absence (e.g. Yasha the character was babysitting while Ashley the 
actor was absent) and on some occasions the other players made decisions 
for the character on the player's, and kept the character in play.

```{r load-data}
# attendance_data <- 
#   read.csv("data/Running Times - Wildemount - WM CR Attendance.csv",
#            stringsAsFactors = FALSE)

attendance_data <- 
  read.csv("data/Running Times - Wildemount - WM CR Attendance 20.02.07.csv",
           stringsAsFactors = FALSE)

attendance_data$Episode.Num <- 
  as.numeric(gsub("C2E", "", attendance_data$Episode))

attendance_data$Date <- 
  as.Date(attendance_data$Air.Date,
                     format = "%m/%d/%y")

attendance_data_long <-
  attendance_data %>%
  tidyr::pivot_longer(cols = c("Laura", 
                               "Liam", 
                               "Marisha", 
                               "Taliesin", 
                               "Travis", 
                               "Sam", 
                               "Ashley"),
                      names_to = "Actor") %>%
  as_tibble() %>%
  dplyr::mutate(value = as.factor(value))
```


```{r high-level-att-heatmap}
# colors <- structure(c("black", "green"), names = c("0", "1"))
# Heatmap(as.matrix(attendance_data[ , 3:9]),
#         col = colors,
#         name = "Critical Role Attendance",
#         column_title = "Actor",
#         column_title_side = "bottom", 
#         cluster_rows = FALSE,
#         column_dend_side = "bottom")
# 
# h1 <- Heatmap(as.matrix(attendance_data[ , 3]),
#         col = mighty_nein_dark_to_light$Travis,
#         name = "Critical Role Attendance",
#         column_title = "Fjord",
#         column_title_side = "bottom", 
#         cluster_rows = FALSE,
#         column_dend_side = "bottom")
# 
# h2 <- Heatmap(as.matrix(attendance_data[ , 4]),
#         col = mighty_nein_dark_to_light$Marisha,
#         name = "Critical Role Attendance",
#         column_title = "Beau",
#         column_title_side = "bottom", 
#         cluster_rows = FALSE,
#         column_dend_side = "bottom")

g1 <- 
  ggplot2::ggplot(data = attendance_data_long %>%
                        dplyr::filter(Actor == "Travis"),
                               aes(x = Episode.Num,
                                   y = Actor,
                                   fill = factor(value))) +
  ggplot2::geom_tile() +
  scale_color_manual(values = as.character(mighty_nein_dark_to_light$Travis),
                     aesthetics = c("colour", "fill"))

g2 <- 
  ggplot2::ggplot(data = attendance_data_long %>%
                        dplyr::filter(Actor == "Marisha"),
                               aes(x = Episode.Num,
                                   y = Actor,
                                   fill = factor(value))) +
  ggplot2::geom_tile() +
  scale_color_manual(values = c("0" = as.character(mighty_nein_dark_to_light$Marisha[1]),
                                "1" = as.character(mighty_nein_dark_to_light$Marisha[2])),
                     aesthetics = c("colour", "fill"))

multiplot(g1, g2)
  

multiplot(h1, h2)

attendance_data_long_with_factor <-
  attendance_data_long %>%
  dplyr::mutate(value_factor = ifelse(test = Episode.Num > 27 & Actor == "Taliesin",
                                      yes = paste0(Actor, as.numeric(value) + 2),
                                      no = paste0(Actor, value))) %>%
  dplyr::mutate(value_factor = factor(value_factor),
                Actor = factor(Actor,
                               levels = c("Ashley",
                                          "Taliesin",
                                          "Laura",
                                          "Sam",
                                          "Liam",
                                          "Marisha",
                                          "Travis"))) 
  


g <- ggplot2::ggplot(data = attendance_data_long_with_factor,
                               aes(x = Episode.Num,
                                   y = Actor,
                                   fill = factor(value_factor))) +
  ggplot2::geom_tile() +
  theme_hc() +
  #theme_clean() +
  # theme_few () +
  scale_color_manual(values = discrete_mighty_nein_colors,
                     aesthetics = c("fill")) +
  scale_x_continuous(breaks = seq(1, 94, 1),
                     labels = paste0("Ep. ", seq(1, 94, 1))) +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 5,
                                   angle = 90),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

cvd_grid(g)
g

## TODO might want to use a charater matrix instead, as this
## will suppress clustering. Depends on what I want to do with
## the data. Maybe clustering of attendance behavior is interesting.

## TODO I want to rotate this so the actors are on the y axis
## and so that the order is the order they sit in on the show
```

```{r load-potion-data}
# potion_data <- read.csv("data/Potions Consumed - Wildemount.csv",
#                         stringsAsFactors = FALSE)

potion_data <- read.csv("data/Potions Consumed - Wildemount - Wildemount Potions 20.12.20.csv",
                        stringsAsFactors = FALSE)

potion_data <-
  potion_data %>%
  filter(! (Administered.To %in% c("Shakäste", "Pumat Sol")))

potion_from <- 
  potion_data %>%
  distinct(Administered.by) %>%
  rename(label = Administered.by)

potion_to <-
  potion_data %>%
  distinct(Administered.To) %>%
  rename(label = Administered.To)

potion_nodes <- 
  full_join(potion_from, potion_to, by = "label") %>%
  rowid_to_column("id")
  
potion_edge_list <- 
  potion_data %>%
  mutate(self_directed = as.numeric(Administered.by == Administered.To)) %>%
  group_by(Administered.by, Administered.To, Potion, self_directed) %>%
  summarize(weight = n()) %>%
  ungroup() %>%
  left_join(potion_nodes, by = c("Administered.by" = "label")) %>% 
  rename(from = id) %>%
  left_join(potion_nodes, by = c("Administered.To" = "label")) %>% 
  rename(to = id) %>%
  select(from, to, weight, Potion, self_directed) %>%
  mutate(weight = as.integer(weight))

agg_potion_edge_list <- 
  potion_data %>%
  group_by(Administered.by, Administered.To) %>%
  summarize(weight = n()) %>%
  ungroup() %>%
  left_join(potion_nodes, by = c("Administered.by" = "label")) %>% 
  rename(from = id) %>%
  left_join(potion_nodes, by = c("Administered.To" = "label")) %>% 
  rename(to = id) %>%
  select(from, to, weight)
```

```{r graph-potions}
# potion_network <- network(potion_edge_list, 
#                           vertex.attr = potion_nodes, 
#                           matrix.type = "edgelist", 
#                           ignore.eval = FALSE)
# plot(potion_network)

#potion_igraph <- graph_from_data_frame(d = potion_edge_list, vertices = potion_nodes, directed = TRUE)

#plot(potion_igraph, layout = layout_with_graphopt, edge.arrow.size = 0.5)

potions_tidy <- tbl_graph(nodes = potion_nodes, edges = potion_edge_list, directed = TRUE)

potions_tidy %>%
  activate(edges) %>%
  arrange(desc(weight))


agg_potions_tidy <- 
  tbl_graph(nodes = potion_nodes, 
            edges = agg_potion_edge_list, 
            directed = TRUE)

agg_potions_tidy %>%
  activate(edges) %>%
  arrange(desc(weight))


# ggraph(potions_tidy, layout = "linear") +
#   geom_edge_loop(aes(x = to + 0.1 - 0.1 * as.numeric(factor(Potion,
#                                                             levels = c("Regular Healing",
#                                                                        "Greater Healing",
#                                                                        "Superior Healing"))),
#                      y = 0,
#                      width = weight, 
#                      color = Potion,
#                      strength = 0.5,
#                      direction = 20,
#                      span = 90),
#                  arrow = arrow(length = unit(2, 'mm')),
#                  end_cap = circle(3, 'mm')) +
#   geom_edge_fan(aes(width = weight, color = Potion), alpha = 0.8,
#                  arrow = arrow(length = unit(2, 'mm')),
#                  end_cap = circle(3, 'mm')) +
#   scale_edge_width(range = c(0.2, 1)) +
#   geom_node_text(aes(label = label)) +
#   labs(edge_width = "Number of Potions Given") +
#   theme_graph()




# ggraph(potions_tidy, layout = "linear") +
  # geom_edge_loop(aes(x = to + 0.1 - 0.1 * as.numeric(factor(Potion,
  #                                                           levels = c("Regular Healing",
  #                                                                      "Greater Healing",
  #                                                                      "Superior Healing"))),
  #                    width = weight,
  #                    color = Potion),
  #                arrow = arrow(length = unit(3, 'mm')),
  #                end_cap = circle(2, 'mm'),
  #                alpha = 0.5) +
  # geom_edge_arc(aes(width = weight,
  #                   color = Potion,
  #                   xend = to + 0.2 - 0.2 * as.numeric(factor(Potion ,
  #                                                             levels = c("Regular Healing",
  #                                                                      "Greater Healing",
  #                                                                      "Superior Healing"))),
  #                   ),
  #                arrow = arrow(length = unit(3, 'mm')), ## TODO need to add var to track self directedness and remove those
  #                end_cap = circle(2, 'mm'),
  #               alpha = 0.5) +
  # scale_edge_width(range = c(0.2, 2)) +
  # geom_node_text(aes(label = label)) +
  # labs(edge_width = "Number") +
  # theme_graph()


image_width <- 0.26
image_height <- 0.4
space_between_images_h <- 0.05
space_between_images_v <- 0.25
image_h_offset <- image_width + space_between_images_h
image_v_offset <- image_height + space_between_images_v
image_x <- 7.0
image_y <- -0.8

arrow_y <- image_y - 0.2
arrow_length <- 0.25
arrow_x_end <-  image_x - 0.5
arrow_x_start <-  image_x - 0.5 - arrow_length
arrow_v_offset <- 0.25

legend_y = image_y + 0.1


## this one
## TODO add white box backgrounds to names to prevent overlap
## TODO fix arrow head angle so it matches the line
background_labels <-
  potion_edge_list %>%
  left_join(potion_nodes,
            by = c("from" = "id")) %>%
  mutate(len = nchar(label)) %>%
  mutate(arc_strength = ifelse(abs(from - to) == 1,
                               0.75,
                               1))


potion_img <- readPNG("img/potion.png")
ggraph(agg_potions_tidy, layout = "linear") +
  geom_edge_loop(mapping = aes(span = 90,
                               direction = 45,
                               x = from + 0.15,
                              # y = 0.15,
                               width = weight),
                 arrow = arrow(length = unit(3, 'mm')),
                 end_cap = circle(3, 'mm'),
                 start_cap = circle(3, "mm"),
                 alpha = 0.8,
                 color = "red4") +
  geom_edge_arc(aes(x = ifelse(from == 4, 
                               from,
                               from - 0.15), 
                    xend = ifelse(from == 4,
                                  to - 0.15,
                                  to),
                    width = weight),
                arrow = arrow(length = unit(3, 'mm')),
                end_cap = circle(3, 'mm'),
                start_cap = circle(3, "mm"),
                alpha = 0.8,
                color = "red4") +
  scale_edge_width(range = c(0.2, 2),
                   labels = c("one", "two", "three", "four")) +
  geom_rect(aes(xmin = to - len/18, 
                ymin = -0.1,
                xmax = to + len/16,
                ymax = 0.1), 
            data = background_labels,
            fill = "white") + 
  geom_node_text(aes(label = label)) +
  labs(edge_width = "Number of Potions Given") +
  theme_graph() +
  ggplot2::annotate("text",
           x = image_x + 0.3,
           y = legend_y,
           label = "Number of Potions Given") +
  ggplot2::annotate("segment",
           x = arrow_x_start,
           xend = arrow_x_end,
           y = arrow_y,
           yend = arrow_y,
           colour = "red4",
           size=0.25,
           alpha=0.8) +
  draw_image(potion_img, 
             x = image_x, 
             y = image_y, 
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  ggplot2::annotate("segment",
           x = arrow_x_start,
           xend = arrow_x_end,
           y = arrow_y - arrow_v_offset,
           yend = arrow_y - arrow_v_offset,
           colour = "red4",
           size=0.75,
           alpha=0.8) +
  draw_image(potion_img, 
             x = image_x, 
             y = image_y - space_between_images_v, 
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + image_h_offset, 
             y = image_y - space_between_images_v,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  ggplot2::annotate("segment",
           x = arrow_x_start,
           xend = arrow_x_end,
           y = arrow_y - 2 *  arrow_v_offset,
           yend = arrow_y - 2 * arrow_v_offset,
           colour = "red4",
           size=1.25,
           alpha=0.8) +
  draw_image(potion_img,
             x = image_x, 
             y = image_y - space_between_images_v * 2,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + image_h_offset, 
             y = image_y - space_between_images_v * 2,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + 2 * image_h_offset, 
             y = image_y - space_between_images_v * 2,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  ggplot2::annotate("segment",
           x = arrow_x_start,
           xend = arrow_x_end,
           y = arrow_y - 3 *  arrow_v_offset,
           yend = arrow_y - 3 * arrow_v_offset,
           colour = "red4",
           size=1.75,
           alpha=0.8) +
  draw_image(potion_img, 
             x = image_x, 
             y = image_y - space_between_images_v * 3,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + image_h_offset, 
             y = image_y - space_between_images_v * 3,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + 2 * image_h_offset, 
             y = image_y - space_between_images_v * 3,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + 2 * image_h_offset, 
             y = image_y - space_between_images_v * 3,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  draw_image(potion_img, 
             x = image_x + 3 * image_h_offset, 
             y = image_y - space_between_images_v * 3,
             hjust = 1, 
             vjust = 1, 
             width = image_width, 
             height = image_height) +
  theme(legend.position = "none")
```


```{r kiri-imitations}
kiri_data <- read.csv("data/Kiri Imitations.csv",
                      stringsAsFactors = FALSE)

kiri_data <-
  kiri_data %>%
  mutate(Imitated = str_match(Kiri, "like .*\\)")) %>%
  mutate(Imitated = str_replace(Imitated, "like ", "")) %>%
  mutate(Imitated = str_replace(Imitated, "\\)", "")) %>%
  mutate(Speech = str_match(Kiri, "[[:alpha:]]\\) .*[[:punct:]]$")) %>%
  mutate(Speech = str_replace(Speech, "[[:alpha:]]\\) ", "")) %>%
  mutate(Words = tolower(Speech)) %>%
  mutate(Words = str_replace(Words, "f\\*\\*", "fuck")) %>%
  mutate(Words = str_replace(Words, "[[:punct:]]", "")) %>%
  mutate(Words = str_replace(Words, "\\.", "")) %>%
  mutate(Words = str_replace(Words, "fuck", "f***")) %>%
  mutate(Number = seq(1:nrow(kiri_data)))

kiri_data_m9 <-
  kiri_data %>%
  filter(Imitated %in% c("Beau",
                         "Caleb",
                         "Fjord",
                         "Jester",
                         "Molly",
                         "Nott",
                         "Yasha")) %>%
  mutate(`Kiri Imitation` = Imitated)

kiri_data_m9 <-
  kiri_data_m9 %>%
  mutate(Number = seq(1:nrow(kiri_data_m9)))

ggplot(data = kiri_data_m9,
       aes(x = Number, y = 0, color = `Kiri Imitation`, label = Speech)) +
  labs(col = "Kiri & The M9: Imitations      ") +
  geom_hline(yintercept = 0,
             color = "white",
             size = 0.3) +
  geom_point(aes(x = Number, color = `Kiri Imitation`)) +
  geom_text(aes(x = Number, y = 0.1, label = Speech, angle = 90, hjust = 0), size = 2, family = "Courier") +
  coord_cartesian(ylim = c(-0.5, 8)) +
  scale_color_manual(values = single_mighty_nein_colors) +
  theme_void() +
    theme(legend.position = "bottom") +
  #ggtitle("Kiri Imitations") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(family="Courier"))
```

